---
title: "introduction to my linear regression model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction to my linear regression model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Introduction

Welcome to the `mylinear` package! This package provides a simple implementation of a linear regression model through the `linear_regression()` function. This vignette demonstrates how to use this function, compares it to R's built-in `lm()` function, and benchmarks its performance.

The `linear_regression()` function estimates the linear relationship between a predictor and a response variable, providing an easy way to understand the relationship between two sets of numeric data.

# Example Usage

To begin, let’s see how to use the `linear_regression()` function to fit a linear regression model to some generated data.
```{r}
# Load the MyLinearReg package
library(mylinear)

# Generate some sample data
set.seed(123) # Setting seed for reproducibility
x <- 1:50
y <- 2 * x + rnorm(10, mean = 0, sd = 1)

# Fit the linear regression model
model <- linear_regression(x, y)

# Display the estimated intercept and slope
print(model)
```

# Comparison with R's Built-in lm() Function

To verify the correctness of the linear_regression() function, we will compare its results with R's built-in lm() function, which is widely used for fitting linear models.

```{r}
lm_model <- lm(y ~ x)
lm_coefficients <- coef(lm_model) 
```


```{r}
# Standardize comparisons
intercept_equal <- all.equal(as.numeric(model$intercept), as.numeric(lm_coefficients[1]))
slope_equal <- all.equal(as.numeric(model$slope), as.numeric(lm_coefficients[2]))

# Print results
cat("Intercept comparison:", intercept_equal, "\n")
cat("Slope comparison:", slope_equal, "\n")
```

# Visualization
```{r fig.width=7, fig.height=5}
# Generate predictions for both models
y_pred_mylinear <- model$intercept + model$slope * x  # Predictions from mylinear
y_pred_lm <- predict(lm_model, newdata = data.frame(x = x))  # Predictions from lm()

# Plot with true data and predictions
plot(
  x, y, 
  main = "Linear Regression with Predictions",
  xlab = "Predictor (x)", 
  ylab = "Response (y)", 
  pch = 16, col = "gray", cex = 0.8, # Weaken true data: lighter color and smaller size
  ylim = range(c(y, y_pred_mylinear, y_pred_lm))
)
lines(x, y_pred_mylinear, col = "blue", lwd = 2) # Prediction line from mylinear
lines(x, y_pred_lm, col = "red", lwd = 2, lty = 2) # Prediction line from lm()
legend(
  "topleft", 
  legend = c("True Data", "linear_regression() Prediction", "lm() Prediction"), 
  col = c("gray", "blue", "red"), 
  pch = c(16, NA, NA), 
  lty = c(NA, 1, 2), 
  lwd = c(NA, 2, 2)
)


```
The plot successfully demonstrates that linear_regression() function performs as expected and gives results consistent with R's built in lm() function.



# Benchmarking Performance
Next, let's compare the performance of linear_regression() with lm() to assess efficiency. We will use the bench package for this purpose.

```{r}
library(bench)

# Benchmarking without enforcing result equality
bench::mark(
  mylinear = {
    # Run the custom linear regression but do not return the result
    linear_regression(x, y)
    NULL # Return a placeholder value
  },
  lm = {
    # Run lm() but do not return the result
    lm(y ~ x)
    NULL # Return a placeholder value
  }
)

```


**Speed**:
My customesed linear_regression(x, y) implementation is significantly faster than lm(y ~ x). `linear_regression()` has a median execution time of 4.8 µs, while `lm()` takes 145.8 µs. `linear_regression()` can run approximately 30 times more iterations per second than `lm()`.

**Memory Usage**:
`linear_regression()` allocates 1.31KB in memory, making it much more memory-efficient compared to `lm()`, which allocates 10.02KB.

**Garbage Collection**:
This lower memory usage results in fewer garbage collection triggers, with `linear_regression()` requiring only one garbage collection per 9999 iterations, while `lm()` requires eleven per 3085 iterations.    



# Conclusion

In this vignette, I demonstrated how to use the `linear_regression()` function from the mylinear package to fit a simple linear model. The results were compared to R's built-in `lm()` function using all.equal() to verify correctness. Additionally, performance was benchmarked using bench::mark() to highlight the efficiency of the custom implementation. These comparisons confirm that the `linear_regression()` function is both accurate and highly efficient, making it a suitable alternative for scenarios where performance is critical.
