---
title: "linear_regression() of mylinear package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction to my linear regression model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Introduction

Welcome to the `mylinear` package! This package provides a simple implementation of a linear regression model through the `linear_regression()` function. This vignette demonstrates how to use this function, compares it to R's built-in `lm()` function, and benchmarks its performance.

The `linear_regression()` function estimates the linear relationship between a predictor and a response variable, providing an easy way to understand the relationship between two sets of numeric data.

## Example Usage

To begin, let’s see how to use the `linear_regression()` function to fit a linear regression model to some generated data.
```{r}
# Load the MyLinearReg package
library(mylinear)

# Generate some sample data
set.seed(123) # Setting seed for reproducibility
x <- 1:10
y <- 2 * x + rnorm(10, mean = 0, sd = 1)

# Fit the linear regression model
model <- linear_regression(x, y)

# Display the estimated intercept and slope
print(model)
```

## Comparison with R's Built-in lm() Function

To verify the correctness of the linear_regression() function, we will compare its results with R's built-in lm() function, which is widely used for fitting linear models.

```{r}
lm_model <- lm(y ~ x)
lm_coefficients <- coef(lm_model) 
```


```{r}
# Standardize comparisons
intercept_equal <- all.equal(as.numeric(model$intercept), as.numeric(lm_coefficients[1]))
slope_equal <- all.equal(as.numeric(model$slope), as.numeric(lm_coefficients[2]))

# Print results
cat("Intercept comparison:", intercept_equal, "\n")
cat("Slope comparison:", slope_equal, "\n")
```


## Benchmarking Performance
Next, let's compare the performance of linear_regression() with lm() to assess efficiency. We will use the bench package for this purpose.

```{r}
library(bench)

# Benchmarking without enforcing result equality
bench::mark(
  mylinear = {
    # Run the custom linear regression but do not return the result
    linear_regression(x, y)
    NULL # Return a placeholder value
  },
  lm = {
    # Run lm() but do not return the result
    lm(y ~ x)
    NULL # Return a placeholder value
  }
)

```


**Speed**:
My customesed linear_regression(x, y) implementation is significantly faster than lm(y ~ x). `linear_regression()` has a median execution time of 4.26 µs, while `lm()` takes 138.87 µs. `linear_regression()` can run approximately 32 times more iterations per second than `lm()`.

**Memory Usage**:
`linear_regression()` does not allocate any additional memory, making it much more memory-efficient compared to `lm()`, which allocates 416 bytes per execution.

**Garbage Collection**:
`lm()` triggers garbage collection more frequently, likely due to the memory allocations it performs. This adds overhead to its execution.



## Conclusion

In this vignette, I demonstrated how to use the `linear_regression()` function from the mylinear package to fit a simple linear model. The results were compared to R's built-in `lm()` function using all.equal() to verify correctness. Additionally, performance was benchmarked using bench::mark() to highlight the efficiency of the custom implementation. These comparisons confirm that the `linear_regression()` function is both accurate and highly efficient, making it a suitable alternative for scenarios where performance is critical.
